Sex and Power: sexual dimorphism in trait variability and its eco-evolutionary and statistical implications

### 2020-05-12 scripts/MouseSexDiffApril20_2020.Rmd
Just doing some cleaning and implementing escalc instead of Al's functions

```{r PopStatsFunction}
calculate_population_stats <- function(mydata, min_individuals = 5) {
  mydata %>%
    group_by(population, strain_name, production_center, sex) %>%
    summarise(
      trait = parameter_name[1],
      x_bar = mean(data_point),
      x_sd = sd(data_point),
      n_ind = n()) %>%
    ungroup() %>%
    filter(n_ind > min_individuals) %>%
    # Check both sexes present & filter those missing
    group_by(population) %>%
    mutate(n_sex = n_distinct(sex)) %>%
    ungroup() %>%
    filter(n_sex == 2) %>%
    select(-n_sex) %>%
    arrange(production_center, strain_name, population, sex)
}
```

@@

```{r EffectSizeFunction}
create_meta_analysis_effect_sizes <- function(mydata) {
  i <- seq(1, nrow(mydata), by = 2)
  input <- data.frame(
    n1i = mydata$n_ind[i],
    n2i = mydata$n_ind[i + 1],
    x1i = mydata$x_bar[i],
    x2i = mydata$x_bar[i + 1],
    sd1i = mydata$x_sd[i],
    sd2i = mydata$x_sd[i + 1]
  )

  mydata[i, ] %>%
    select(strain_name, production_center, trait) %>%
    mutate(
      effect_size_CVR = calculate_lnCVR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      sample_variance_CVR = calculate_var_lnCVR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      effect_size_VR = calculate_lnVR(CSD = input$sd1i, CN = input$n1i, ESD = input$sd2i, EN = input$n2i),
      sample_variance_VR = calculate_var_lnVR(CN = input$n1i, EN = input$n2i),
      effect_size_RR = calculate_lnRR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      sample_variance_RR = calculate_var_lnRR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      err = as.factor(seq_len(n()))
    )
}
```

@@

# Create dataframe to store results
results_alltraits_grouping <- 
    tibble(id = 1:n, lnCVR=0, lnCVR_lower=0, lnCVR_upper=0, 
           lnCVR_se=0, lnVR=0, lnVR_lower=0, lnVR_upper=0, 
           lnVR_se=0, lnRR=0, lnRR_lower=0, lnRR_upper=0, lnRR_se=0, sampleSize=0, trait=0)

for (t in 1:n) {
  tryCatch(
    {
      results <- data %>% 
        data_subset_parameterid_individual_by_age(t) %>%
        calculate_population_stats() %>%
        create_meta_analysis_effect_sizes()

@@

      results_alltraits_grouping[t, 2:14] <- c(f(cvr), f(cv), f(means), means$k)
      results_alltraits_grouping[t, 15] <- unique(results$trait)
    },
    error = function(e) {
      cat("ERROR :", t, conditionMessage(e), "\n")
    }
  )
}

>>>>

```{r PopStatsFunction}
calculate_population_stats <- function(mydata, min_individuals = 5) {
  mydata %>%
    group_by(population, 
             strain_name, 
             production_center, 
             sex) %>%
    summarise(trait = parameter_name[1],
              x_bar = mean(data_point),
              x_sd = sd(data_point),
      n_ind = n()) %>%
    ungroup() %>%
    filter(n_ind > min_individuals) %>%
    # Check both sexes present & filter those missing
    group_by(population) %>%
    mutate(n_sex = n_distinct(sex)) %>%
    ungroup() %>%
    filter(n_sex == 2) %>%
    select(-n_sex) %>%
    arrange(production_center, strain_name, population, sex)
}

multi_spread <- function(df, key, value) {
    # quote key
    keyq <- rlang::enquo(key)
    # break value vector into quotes
    valueq <- rlang::enquo(value)
    s <- rlang::quos(!!valueq)
    df %>% gather(variable, value, !!!s) %>%
        unite(temp, !!keyq, variable) %>%
        spread(temp, value)
}

# Will calculate population stats and exclude data that is irrelevant and not useful, it will also just convert the data straight away to wide format as opposed to long, which I think is making things more complicated when effect sizes are calculated.

calculate_population_stats_new <- function(mydata, min_individuals = 5){
    mydata %>% 
    group_by(population,
             production_center,  
             strain_name,  
             sex) %>%
    summarise(trait  = parameter_name[1],
              x_bar  = mean(data_point, na.rm = TRUE),
              x_sd   =   sd(data_point, na.rm = TRUE),
              n_ind  = n()) %>%
    ungroup() %>%
    filter(n_ind > min_individuals) %>%   # NOTE here that you are excluding data with 5 exactly....
    group_by(population) %>%
    filter(all(c("male", "female") %in% sex)) %>% # Removes any entries where only 1 sex is present
    multi_spread(sex, c(x_bar, x_sd, n_ind))      # Spreads the data out into wide format, so sexes are beside each other which is more typical of meta-analytic data
}

@@

```{r EffectSizeFunction}

#This function should mostly replace the entire function below, but I am still not quite sure what is happening in places. Adding all effect sizes together (maintaining the naming), err seems to be added to each trait? That needs to be added.It will make use of escalc which has the correct equations for effect size calculations.
create_meta_analysis_effect_sizes_new <- function(mydata, measure = c("CVR", "VR", "ROM")){
  for(i in 1:length(measure)){
    mydata <- metafor::escalc(m1i  = female_x_bar, 
                              m2i  = male_x_bar, 
                              sd1i = female_x_sd, 
                              sd2i = male_x_sd, 
                              n1i  = female_n_ind, 
                              n2i  = male_n_ind, 
                              data = mydata, 
                           measure = measure[i], 
                         var.names = c(paste0("effect_size_", measure[i]), 
                                      paste0("sample_variance_", measure[i])),
                            append = TRUE)
    }
    return(mydata)
}

## Can delete all the below ###### The above function does same thing, but uses simplified dataset and also reorganises it so no need for input dataframe. This also means we can remove the next section with Al's function as escalc has these built in anyway.
create_meta_analysis_effect_sizes <- function(mydata) {
  i <- seq(1, nrow(mydata), by = 2)
  input <- data.frame(
    n1i = mydata$n_ind[i],
    n2i = mydata$n_ind[i + 1],
    x1i = mydata$x_bar[i],
    x2i = mydata$x_bar[i + 1],
    sd1i = mydata$x_sd[i],
    sd2i = mydata$x_sd[i + 1]
  )

  mydata[i, ] %>%
    select(strain_name, production_center, trait) %>%
    mutate(
      effect_size_CVR = calculate_lnCVR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      sample_variance_CVR = calculate_var_lnCVR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      effect_size_VR = calculate_lnVR(CSD = input$sd1i, CN = input$n1i, ESD = input$sd2i, EN = input$n2i),
      sample_variance_VR = calculate_var_lnVR(CN = input$n1i, EN = input$n2i),
      effect_size_RR = calculate_lnRR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      sample_variance_RR = calculate_var_lnRR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      err = as.factor(seq_len(n()))
    )

}
```

@@

# Create dataframe to store results
results_alltraits_grouping <- 
    tibble(id = 1:n, 
           lnCVR=0, lnCVR_lower=0, lnCVR_upper=0,lnCVR_se=0, 
            lnVR=0,  lnVR_lower=0,  lnVR_upper=0, lnVR_se=0, 
            lnRR=0,  lnRR_lower=0,  lnRR_upper=0, lnRR_se=0, 
           sampleSize=0, trait=0)

for (t in 1:n) {
  tryCatch(
    {

    #NOTE: results and results 2 should now be equivalent and produce the same thing except that results 2 has the columns added in with raw effect sizes, which I think is wise if we can keep them. Otherwise, the naming should be the same. I've checked quite a few t's and all looks the same. results 2 just has more columns.
      results <- data.frame(data %>% 
        data_subset_parameterid_individual_by_age(t) %>%
        calculate_population_stats() %>%
        create_meta_analysis_effect_sizes())

      results2 <- data %>%
       data_subset_parameterid_individual_by_age(t) %>%
       calculate_population_stats_new() %>%
       create_meta_analysis_effect_sizes_new() %>%
       mutate(err = seq_len(n()))

@@

      results_alltraits_grouping[t, 2:14] <- c(f(cvr), f(cv), f(means), means$k)
      results_alltraits_grouping[t, 15] <- unique(results$trait)
    },
    error = function(e) {
      cat("ERROR :", t, conditionMessage(e), "\n")
    }
  )
}
```
